Incremental Forward Reasoning for White-Box~Proof~Search

### 2025-02-25 paper.tex
fix big O and broken indices

More specifically, we first match the type of the last slot, $Aₙ$, with each hypothesis $h : B$ (using a discrimination tree index to quickly determine potentially matching hypotheses).
If successful, this results in a substitution $σ₁$ containing all variables that occur in $Aₙ$.
We then match $A_{ₙ-1}[σ₁]$ against the type of each hypothesis in the context, yielding a substitution $σ₂$, and proceed with $A_{ₙ-2}[σ₂]$, etc., until we obtain a complete match for $r$.
If a premise $A_{ₙ-i}[σᵢ]$ does not match any hypothesis type, we backtrack and consider other potential matching hypotheses for $A_{ₙ-i+1}$.
These may lead to a different substitution $τ$ and therefore to different matching hypotheses for $A_{ₙ-i}[τ]$.
For each complete match $m$ with substitution $σ$ found during this process, we add a new hypothesis of type $B[σ]$, where $B$ is the conclusion of $r$.

@@

These results are unsurprising, given the complexity of both algorithms on this benchmark.
The naive algorithm goes over the context, containing initially $n$ hypotheses, and compares all hypotheses pairwise.
This is repeated approximately for each of the $n(n-1)/2$ added hypotheses, so the overall complexity is $O(n^4)$.
By contrast, the incremental algorithm considers each hypothesis only once, matching it against both slots and then comparing it with at most $n$ other hypotheses, so the complexity is $O(n^{2})$.

@@

Still, Fig.~\ref{fig:indep} shows that the incremental approach outperforms the naive one for $n ≥ 4$.
This is because incrementality helps even in this scenario.
Whenever a hypothesis is added (and hence a new subgoal created), the naive algorithm re-runs all the rules it had already run on the previous subgoal, realising only after the fact that the rules were already applied.
The incremental algorithm avoids this source of redundancy.
Accordingly, the naive algorithm has complexity $O(n^{3})$ on this benchmark, versus $O(n)$ for the incremental algorithm.

>>>>

More specifically, we first match the type of the last slot, $Aₙ$, with each hypothesis $h : B$ (using a discrimination tree index to quickly determine potentially matching hypotheses).
If successful, this results in a substitution $σ₁$ containing all variables that occur in $Aₙ$.
We then match $A_{n-1}[σ₁]$ against the type of each hypothesis in the context, yielding a substitution $σ₂$, and proceed with $A_{n-2}[σ₂]$, etc., until we obtain a complete match for $r$.
If a premise $A_{n-i}[σᵢ]$ does not match any hypothesis type, we backtrack and consider other potential matching hypotheses for $A_{n-i+1}$.
These may lead to a different substitution $τ$ and therefore to different matching hypotheses for $A_{n-i}[τ]$.
For each complete match $m$ with substitution $σ$ found during this process, we add a new hypothesis of type $B[σ]$, where $B$ is the conclusion of $r$.

@@

These results are unsurprising, given the complexity of both algorithms on this benchmark.
The naive algorithm goes over the context, containing initially $n$ hypotheses, and compares all hypotheses pairwise.
This is repeated approximately for each of the $n(n-1)/2$ added hypotheses, so the overall number of unifications performed is of the order of $n^4$.
By contrast, the incremental algorithm considers each hypothesis only once, matching it against both slots and then comparing it with at most $n$ other hypotheses. In this case, the number of unifications performed is of the order of $n^{2}$.

@@

Still, Fig.~\ref{fig:indep} shows that the incremental approach outperforms the naive one for $n ≥ 4$.
This is because incrementality helps even in this scenario.
Whenever a hypothesis is added (and hence a new subgoal created), the naive algorithm re-runs all the rules it had already run on the previous subgoal, realising only after the fact that the rules were already applied.
The incremental algorithm avoids this source of redundancy.
Accordingly, the naive algorithm performs $n^{3}$ unifications on this benchmark, versus $n$ for the incremental algorithm.
